Multidimensional analysis and detection of informative features in human white matter

### 2021-01-25 reports/manuscript/sections/introduction.tex
Mostly work in results section.
Update reports/manuscript/sections/introduction.tex
Co-authored-by: Adam Richie-Halford <richford@users.noreply.github.com>

\begin{figure}[!h]
    \vspace{0.25cm}
    \rlap{
        \hspace{0.5em}
        \includegraphics[width=0.98\columnwidth]{methods-brains.pdf}
    }
    \vspace{-9.75cm}
    \includegraphics[width=\columnwidth]{method-quad.pdf}
    {\phantomsubcaption\label{fig:methods:tractogram}}
    {\phantomsubcaption\label{fig:methods:cst}}
    {\phantomsubcaption\label{fig:methods:tract-profile}}
    {\phantomsubcaption\label{fig:methods:group-structure}}
    \caption{{\bf Tractometry data lineage}
        \label{fig:methods}
        {\bf (a)} Whole brain tractography generates streamlines approximating
        the trajectories of white matter connections.
        {\bf (b)} Tractometry classifies these streamlines into anatomical bundles.
        In this case, we show the left corticospinal tract (CSTL) over a mid-saggital
        anatomical slice.
        {\bf (c)} Tractometry further extracts bundle profiles,
        quantifications of various diffusion metrics along the length of the
        fiber bundle. Here, we show one subject's fractional anisotropy (FA)
        profile for CSTL.
        {\bf (d)} the phenotypical target data and tractometric features can
        be organized into a linear model, $\hat{y} = \mathbf{X}
        \hat{\beta}$. The feature matrix $\mathbf{X}$ is color-coded
        to reveal a natural group structure: the left (orange) group
        contains $k$ features from the CSTL, the middle (green) group
        contains $k$ features from the left cingulum cingulate (CGCL),
        and the right (blue) group
        contains $k$ features from the left arcuate (ARCL). The coefficients in
        $\hat{\beta}$ follow the same natural grouping.
        % Cross-validation:
        % we evaluate model quality using a nested $k$-fold cross validation
        % scheme. At level-0, the input data is decomposed into $k$
        % shuffled groups and optimal hyperparameters are found for the level-0
        % training set.
        % To avoid overfitting, the optimal hyperparameters are themselves
        % evaluated using a cross-validation scheme taking place
        % at level-1 of the decomposition, where each
        % level-0 training set is further decomposed into five shuffled
        % groups. For the ALS and WH data, $k=10$, while for the HBN and
        % Cam-CAN data, $k=5$.
    }
\end{figure}

@@

The Lasso algorithm minimizes the sum of the absolute values of
contributions of each feature \cite{Tibshirani1996-qs}. This
tends to shrink to zero the contributions of many of the features,
providing results that are both accurate and interpretable. When
additional structure is available in the organization of the data,
regularization algorithms can take advantage of this structure. For
example, if the features lend themselves to a natural division into
different groups, the group lasso (GL) can be used to select groups
of features, rather than individual features \cite{Yuan2006-ky}.
The Sparse Group Lasso (SGL) elaborates on this idea by providing
control both of group sparsity, as well as overall sparsity of the
solutions \cite{simon2013sgl}. Because the features measured with
tractomery lend themselves to grouping based on the tracts from which
each measurement is taken, GL and SGL provide a useful tool
for linear model fitting in problems of this form. Here we, first,
develop an implementation of SGL that is well suited to the analysis of
tractometry data and, second, demonstrate the power and flexibility of
this approach by applying it to both classification
and continuous prediction problems.

>>>>

\begin{figure}[!h]
    \vspace{0.25cm}
    \rlap{
        \hspace{0.5em}
        \includegraphics[width=0.98\columnwidth]{methods-brains.pdf}
    }
    \vspace{-9.75cm}
    \includegraphics[width=\columnwidth]{method-quad.pdf}
    {\phantomsubcaption\label{fig:methods:tractogram}}
    {\phantomsubcaption\label{fig:methods:cst}}
    {\phantomsubcaption\label{fig:methods:tract-profile}}
    {\phantomsubcaption\label{fig:methods:group-structure}}
    \caption{{\bf Tractometry data flow}
        \label{fig:methods}
        {\bf (a)} Whole brain tractography generates streamlines approximating
        the trajectories of white matter connections.
        {\bf (b)} Tractometry classifies these streamlines into anatomical bundles.
        In this case, we show the left corticospinal tract (CSTL) over a mid-saggital
        anatomical slice.
        {\bf (c)} Tractometry further extracts bundle profiles,
        quantifications of various diffusion metrics along the length of the
        fiber bundle. Here, we show one subject's fractional anisotropy (FA)
        profile for CSTL.
        {\bf (d)} the phenotypical target data and tractometric features can
        be organized into a linear model, $\hat{y} = \mathbf{X}
        \hat{\beta}$. The feature matrix $\mathbf{X}$ is color-coded
        to reveal a natural group structure: the left (orange) group
        contains $k$ features from the CSTL, the middle (green) group
        contains $k$ features from the left cingulum cingulate (CGCL),
        and the right (blue) group
        contains $k$ features from the left arcuate (ARCL). The coefficients in
        $\hat{\beta}$ follow the same natural grouping.
        % Cross-validation:
        % we evaluate model quality using a nested $k$-fold cross validation
        % scheme. At level-0, the input data is decomposed into $k$
        % shuffled groups and optimal hyperparameters are found for the level-0
        % training set.
        % To avoid overfitting, the optimal hyperparameters are themselves
        % evaluated using a cross-validation scheme taking place
        % at level-1 of the decomposition, where each
        % level-0 training set is further decomposed into five shuffled
        % groups. For the ALS and WH data, $k=10$, while for the HBN and
        % Cam-CAN data, $k=5$.
    }
\end{figure}

@@

The Lasso algorithm minimizes the sum of the absolute values of
contributions of each feature \cite{Tibshirani1996-qs}. This
tends to shrink to zero the contributions of many of the features,
providing results that are both accurate and interpretable. When
additional structure is available in the organization of the data,
regularization algorithms can take advantage of this structure. For
example, if the features lend themselves to a natural division into
different groups, the group lasso (GL) can be used to select groups
of features, rather than individual features \cite{Yuan2006-ky}.
The Sparse Group Lasso (SGL) elaborates on this idea by providing
control both of group sparsity, as well as overall sparsity of the
solutions \cite{simon2013sgl}. Because the features measured with
tractomery lend themselves to grouping based on the tracts from which
each measurement is taken, GL and SGL provide useful tools
for linear model fitting in problems of this form. Here we, first,
develop an implementation of SGL that is well suited to the analysis of
tractometry data and, second, demonstrate the power and flexibility of
this approach by applying it to both classification
and continuous prediction problems.
